Multi-Instance Methods for Partially Supervised Image Segmentation

### 2013-08-28 partial_supervision/paper.tex
Remove MIL theory stuff, start rereading partial supervision part

The task of multi-class segmentation is often split up in a segmentation
and a recognition part. Random forest methods often classify each pixel
separately and segment using predicted classes~\citep{schroff2008object} while SVM-based methods often work on an
over-segmentation of the image, called superpixels~\citep{ladicky2009associative,gonfaus2010harmony}. Superpixels avoid the
computational burden of classifying each pixel separately, but have
two drawbacks:
\begin{enumerate}
    \item A single superpixel does not provide enough context for classification~\citep{fulkerson2009class}.
    \item Segment boundaries are decided on the lowest level by generating
        superpixels. This decision cannot be corrected afterwards~\citep{hanbury2008superpixels}.
\end{enumerate}
In our approach, we work with a set of candidate segments, generated using constrained parametric min-cuts~\citep{carreira2010constrained}.
For each image, these segments are a set of overlapping, object-like regions, which serve
as candidates for object locations.

We formulate weakly supervised multi-class image segmentation as a multi-instance problem, based upon candidate segments.
In multi-instance learning~\citep{dietterich1997solving}, each training example is given as a multi-set of instances, called a bag.
Each instance is represented as a feature vector $x$ and a label $y$.
A bag is labeled positive if it contains at least one positive example, and negative otherwise.
During training, only the labels of the training bags, not of the instances inside the bags, are known.
The goal is to learn a classifier for unseen bags. 
Formally, let $\mathcal{X}$ be the set of instances. To simplify notation, we assume that bags are simply sets, not multi-sets.
Then a bag is an element of the power set $2^\mathcal{X}$ and the task is
to learn a function
\begin{equation} f_{MI} \colon 2^\mathcal{X} \rightarrow \{-1,+1\}.  \end{equation}
Training examples are tuples $(X_i,y_i)$ of bags $X_i \subset \mathcal{X}$ and labels $y_i \in \{-1,+1\}$.
It is often assumed that the $f_{MI}$ function stems from a so-called underlying concept, given by an (unknown) function
$f_{I} \colon \mathcal{X} \rightarrow \{-1,+1\}$, with 
\begin{equation}\eqlabel{multi_instance}
f_{MI}(X)= \max_{x \in X} f_{I}(x).
\end{equation}

Sometimes, the goal of finding $f_{MI}$ is extended to finding labels not only
for new bags, but also for all the separate instances within a bag
\citep{liconvex2010,zha2008joint}, that is finding $f_{I}$.\\ 
Even though finding $f_{I}$ is sometimes included in the task statement, there
has been very little work that actually reported accuracy on instance label
prediction. Part of the reason for this might be that for many of the datasets
used in multi-instance learning no ground truth exists.

@@

\subsection{Multi-Instance Methods}
Multi-instance learning was formally introduced in \citet{dietterich1997solving}.
Since then, many algorithms were proposed to solve the multi-instance learning problem
using many different approaches \citep{andrews2003support,gaertner2002multi,zhou2009multi,li2009convex,zhang2002dd,mangasarian2008multiple,leistner2010miforests,chen2006miles}.
We will discuss only those that are relevant to the present treatment.

\citet{gaertner2002multi} introduced the concept of a multi-instance
kernel on bags, defined in terms of a kernel on instances. 
The basic principle of multi-instance kernel is similar to a soft-max over instances in
each bag. This can be viewed as approximating the kernel value of the ``closest pair'' given by two bags. They show that the multi-instance kernel
is able to separate bags if and only if the original kernel on instances is able to separate the underlying concepts.
The method of \citet{gaertner2002multi} has a particular appeal in that it transforms a multi-instance problem into a standard
classification problem by changing the kernel. The downside of this approach is that it does
not directly label instances, only bags.
 %and that it assumes that instances inside a bag have i.i.d.
%labels.

\citet{zhou2009multi} explicitly address non-i.i.d.\  labels, leading to an algorithm that can take
advantage of correlations inside bags. Computational costs of their algorithm does not scale well with
the number of instances, although a heuristic algorithm is proposed to overcome this restriction.
\citet{zhou2009multi} demonstrated only a slight advantage of their algorithm over the MI-kernel of \citet{gaertner2002multi}, so
we use the MI-kernel for better scalability.

@@

While semantic segmentation is closely related to multi-class image segmentation, there are important
distinctions: In semantic segmentation, each pixel has a semantic annotation, also containing ``background''
classes like ``sky'', ``grass'' and ``water''. In multi-class image segmentation, the focus is on
objects, with possibly large parts of the image being labeled as unspecific ``background''.
The unspecific background class contains much more clutter than for example ``grass'' and is therefore
much harder to model.
This makes disseminating the interesting part in multi-class object recognition challenging, since it
is not necessary possible to identify non-object regions easily.

\section{Multi-Instance Kernels for Image Segmentation}
\subsection{Constraint Parametric Min Cuts (CPMC)}
To generate proposal segments, we use the Constraint Parametric Min Cut (CPCM) method from \citet{carreira2010constrained}. In CPMC, initial
segments are constructed graph cuts on the pixel grid. The energy function for these cuts uses pixel
color and the response of the global probability of boundary (gPb) detector~\citep{maire2008using}. As much as ten thousand initial segments are generated from foreground
and background seeds. A fast rejection based on segment size and ratio cut~\citep{wang2003image} reduced these to about 2000 overlapping
segments per image. Then, the segments are ranked according to a measure of object-likeness that is based on region and Gestalt properties.
This ranking is computed using an SVR model~\citep{carreira2010constrained}, which is available online.
For computing the global probability of boundary (gPb), we used the CUDA implementation
of \citet{catanzaro2009efficient}, which provides a speedup of two orders of magnitude over the original implementation.

\subsection{Multi-Instance Learning using MI-Kernels}
Since scalability is very important in real-world computer vision
applications, and natural images might need hundreds of segments to
account for all possible object boundaries, we use the efficient
multi-instance kernels~\citep{gaertner2002multi}.
Multi-instance kernels are a form of set kernels that transform a kernel
on instance level to a kernel on bag level.
We reduce the multi-instance multi-class problem to a multi-instance problem by using the one-vs-all approach.

With $k_I$ denoting a kernel on instances $x,x' \in \mathcal{X}$, we define the corresponding multi-instance kernel $k_{MI}$ on bags $X,X' \in 2^\mathcal{X}$
as
\begin{equation}\eqlabel{mi_kernel}
k_{MI}(X,X') := \sum_{x \in X, x' \in X'} k^p(x,x'),
\end{equation}
where $p \in \mathbb{N}$ is a parameter~\citep{gaertner2002multi}. As we use the RBF-kernel $k_{rbf}$ as kernel on $\mathcal{X}$ and powers of RBF-kernels
are again RBF-kernels, we will not consider $p$ in the following.

@@

\subsection{Segment Features}
To describe single segments, we make use of densely computed SIFT~\citep{lowe2004distinctive} and ColorSIFT~\citep{van2009evaluating} features, from which
we compute bag of visual word histograms. Additionally, we use histograms of oriented gradients~\citep{dalal2005histograms} on the segments.
We use RBF-kernels for all of the features,
constructing one MI-kernel per feature. These are then combined using multiple kernel
learning to produce a single kernel matrix. This kernel matrix can then be used for all classes,
making classification particularly efficient.

\subsection{Combining Segments}\seclabel{combiningsegments}
The framework described above yields an image-level and a segment-level classification.
In our setup, each segment might be given multiple labels. %TODO explain!
To obtain a pixel-level object-class segmentation, we have to combine these. 
When building the segmentation for a given image, we only consider classes whose presence was predicted on image level.
Since we do not make use of the ground truth segmentation during training, we cannot learn an optimal combination as 
in \citet{li2010object} but perform a simple majority vote instead.
We merge segments into pixel-level class labels by setting
the label $y_x$ of a pixel $x$ according to
\begin{equation}
    y_x = \argmax_{y \in Y} \#\{S_i | p \in S_i \land y_{S_i}=y \},
\end{equation}
where $Y= \{$car, bike, person$ \}$, $S_i$ enumerates all segments within an %FIXME we didn't describe the dataset yet!
image and $y_{S_i}$ is the label of segment $S_i$. In words: each pixel is
assigned the class with the highest number of class segments containing it.
This simple heuristic yields good results in practice.

\begin{figure}[tbp]
	\begin{center}
        \includegraphics[width=\linewidth]{images/scheme-crop.pdf}
	\end{center}
        \caption{Overview of our method. See text for details.}
	\figlabel{results}
\end{figure}
\section{Experiments}
\subsection{Instance-Level Predictions using Multi-Instance Kernels}
To assess the validity of instance-level predictions using multi-instance kernels, we transform $f_{I}$
back to an instance-level classifier, using the multi-instance learning
assumption (\eqref{multi_instance}). We refer to these instance-based MIK predictions
as MIK-instance. In all experiments, the parameters of the MI-Kernel
and SVM are adjusted using MIK and then used with both MIK and MIK-instance.
This facilitates very fast parameter scans since MIK is very efficient to
compute. Note that we cannot adjust parameters using instance prediction error, as we assume
no instance labels to be known.

@@

\end{table}
We compared the performance of MIK, MIK-instance and state-of-the-art MI
methods on the Musk benchmark datasets~\citep{dietterich1997solving}, see
\Tabref{muskacc}. Results were obtained using 10-fold
cross-validation. While the computational complexity of MIK-instance is
very low compared to the other methods, it achieves competitive results.
Using instance-level labels results in a slight loss of accuracy of
MIK-instances, compared to MIK\@. This small degradation of performance is quite surprising, since the model was
not trained to provide any instance-level labels.
i
For multi-class image segmentation, it is beneficial to have a low witness
rate, that is only a few instances are assumed to be positive in a positive
bag. Since an object might not be very prominent in an image, only a fraction
of segments might correspond to the object.
\Tabref{musk-witness} compares the witness rates of MIK-instance,
miSVM~\citep{andrews2003support} and SVR-SVM~\citep{liconvex2010} on the Musk
datasets. MIK-instance is able to achieve similar accuracy
with much less witnesses than the other methods.  Note that Musk1 consists of
very small bags while Musk2 contains significantly larger bags, more similar to the
image/segment setup.
\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccc}
    \toprule
    & \multicolumn2c{Musk1}  & \multicolumn2c{Musk2}  \\
                &accuracy & witness-rate & accuracy & witness-rate  \\
    \cmidrule(rl){2-3}
    \cmidrule(rl){4-5}
    mi-SVM      & 87.4          & 100\%               &  83.6          & 83.9\%\\
    SVM-SVR     & 87.9          & 100\%               &  85.4          & 89.5\%\\
    MIK-instance& 88.0          & 99\%                &  85.2          & 62.3\%\\
    \bottomrule
    \end{tabularx}
    \caption{MIL algorithms and the empirical witness rates of the
        classifiers.  \tablabel{musk-witness}}
\end{table}

@@

\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}p{208pt}ccc}
    \toprule
                & car & bike & person \\
    \cmidrule(l){2-2}
    \cmidrule(l){3-3}
    \cmidrule(l){4-4}
        This work&   0.30&  0.45&  0.43 \\
        Best strongly supervised approaches~\citep{fulkerson2009class,schulz2011}&   0.72&  0.72&  0.66 \\
    \bottomrule
    \end{tabularx}
    \caption{Pixel-level accuracy on the Graz-02 dataset.
        \tablabel{tab:graz}}
\end{table}
Per-class pixel accuracies are reported in \Tabref{tab:graz}; some qualitative results are shown in \figref{results}. The overall
accuracy on images labels, which is the task that was actually trained, is $90\%$.
While the performance of our multiple-instance based approach is far
from current methods that use pixel-level annotations, whose pixel-level accuracy is around $70\%$~\citep{fulkerson2009class,schulz2011} on pixel-level,
it can serve as a baseline for research on weakly supervised methods for image segmentation.

>>>>

%The task of multi-class segmentation is often split up in a segmentation
%and a recognition part. Random forest methods often classify each pixel
%separately and segment using predicted classes~\citep{schroff2008object} while SVM-based methods often work on an
%over-segmentation of the image, called superpixels~\citep{ladicky2009associative,gonfaus2010harmony}. Superpixels avoid the
%computational burden of classifying each pixel separately, but have
%two drawbacks:
%\begin{enumerate}
    %\item A single superpixel does not provide enough context for classification~\citep{fulkerson2009class}.
    %\item Segment boundaries are decided on the lowest level by generating
        %superpixels. This decision cannot be corrected afterwards~\citep{hanbury2008superpixels}.
%\end{enumerate}
In our approach, we work with a set of candidate segments, generated using constrained parametric min-cuts~\citep{carreira2010constrained}.
For each image, these segments are a set of overlapping, object-like regions, which serve
as candidates for object locations.

We formulate weakly supervised multi-class image segmentation as a multi-instance problem, based on candidate segments.
In multi-instance learning~\citep{dietterich1997solving}, each training example is given as a multi-set of instances, called a bag.
Each instance is represented as a feature vector $x$ and a label $y$.
A bag is labeled positive if it contains at least one positive example, and negative otherwise.
During training, only the labels of the training bags, not of the instances inside the bags, are known.
The goal is to learn a classifier for unseen bags. 
Formally, let $\mathcal{X}$ be the set of instances. To simplify notation, we assume that bags are simply sets, not multi-sets.
Then a bag is an element of the power set $2^\mathcal{X}$ and the task is
to learn a function
\begin{equation} f_{MI} \colon 2^\mathcal{X} \rightarrow \{-1,+1\}.  \end{equation}
Training examples are tuples $(X_i,y_i)$ of bags $X_i \subset \mathcal{X}$ and labels $y_i \in \{-1,+1\}$.
It is often assumed that the $f_{MI}$ function stems from a so-called underlying concept, given by an (unknown) function
$f_{I} \colon \mathcal{X} \rightarrow \{-1,+1\}$, with 
\begin{equation}\eqlabel{multi_instance}
f_{MI}(X)= \max_{x \in X}\ f_{I}(x).
\end{equation}

Sometimes, the goal of finding $f_{MI}$ is extended to finding labels not only
for new bags, but also for all the separate instances within a bag
\citep{liconvex2010,zha2008joint}, that is finding $f_{I}$.\\ 
Even though finding $f_{I}$ is sometimes included in the task statement, there
has been very little work that actually reported accuracy on instance label
prediction. Part of the reason for this might be that for many of the datasets
used in multi-instance learning no ground truth for instances exists.

@@

\subsection{Multi-Instance Methods}
Multi-instance learning was formally introduced in \citet{dietterich1997solving}.
Since then, many algorithms were proposed to solve the multi-instance learning problem
\citep{andrews2003support,gaertner2002multi,zhou2009multi,li2009convex,zhang2002dd,mangasarian2008multiple,leistner2010miforests,chen2006miles}.
We will discuss only those that are relevant to the present treatment.

\citet{gaertner2002multi} introduced the concept of a multi-instance
kernel on bags, defined in terms of a kernel on instances. 
The basic principle of multi-instance kernel is similar to a soft-max over instances in
each bag. This can be viewed as approximating the kernel value of the ``closest pair'' given by two bags. They show that the multi-instance kernel
is able to separate bags if and only if the original kernel on instances is able to separate the underlying concepts.
The method of \citet{gaertner2002multi} has a particular appeal in that it transforms a multi-instance problem into a standard
classification problem by using an appropriate kernel. The downside of this approach is that it does
not directly label instances, only bags.
 %and that it assumes that instances inside a bag have i.i.d.
%labels.

\citet{zhou2009multi} explicitly address the fact that instances are not
indpendent within a bag , leading to an algorithm that can take advantage of
possible correlations. Computational costs of their algorithm does not scale
well with the number of instances, although a heuristic algorithm is proposed
to overcome this restriction. 
%FIXME where does this sentence go?
\citet{zhou2009multi} demonstrated only a slight
advantage of their algorithm over the MI-kernel of \citet{gaertner2002multi},
so we use the MI-kernel for better scalability.

@@

While semantic segmentation is closely related to multi-class image
segmentation, there are important distinctions: In semantic segmentation, each
pixel has a semantic annotation, also containing non-object ``stuff'' classes like
``sky'', ``grass'' and ``water''. In multi-class image segmentation, the focus
is on objects, with possibly large parts of the image being labeled as
unspecific ``background''.  The unspecific background class contains much more
clutter than for example ``grass'' and is therefore much harder to model.
This makes disseminating the interesting part in multi-class object recognition
challenging, since it is not necessary possible to identify non-object regions
easily.

\section{Multi-Instance Kernels for Image Segmentation}
\subsection{Constraint Parametric Min Cuts (CPMC)}
% Repetition?
To generate proposal segments, we use the Constraint Parametric Min Cut (CPCM)
method from \citet{carreira2010constrained}. In CPMC, initial segments are
constructed using graph cuts on the pixel grid. The energy function for these cuts
uses pixel color and the response of the global probability of boundary (gPb)
detector~\citep{maire2008using}. As much as ten thousand initial segments are
generated from foreground and background seeds. A fast rejection based on
segment size and ratio cut~\citep{wang2003image} reduces these to about 2000
overlapping segments per image. Then, the segments are ranked according to a
measure of object-likeness that is based on region and Gestalt properties.
This ranking is computed using an SVR model~\citep{carreira2010constrained},
which is available online. 
% TODO implementation detail?
For computing the global probability of boundary
(gPb), we used the CUDA implementation of \citet{catanzaro2009efficient}, which
provides a speedup of two orders of magnitude over the original implementation.

\subsection{Multi-Instance Learning using MI-Kernels}
Since scalability is very important in real-world computer vision applications,
and natural images might need hundreds of segments to account for all possible
object boundaries, we use the efficient multi-instance
kernels~\citep{gaertner2002multi}.  Multi-instance kernels are a form of set
kernels %TODO add reference to earlier work
 that transform a kernel on instance level to a kernel on bag level.  We
reduce the multi-instance multi-class problem to a multi-instance problem by
using the one-vs-all approach.

With $k_I$ denoting a kernel on instances $x,x' \in \mathcal{X}$, we define the
corresponding multi-instance kernel $k_{MI}$ on bags $X,X' \in 2^\mathcal{X}$
as
\begin{equation}\eqlabel{mi_kernel}
k_{MI}(X,X') := \sum_{x \in X, x' \in X'} k^p(x,x'),
\end{equation}
where $p \in \mathbb{N}$ is a free parameter~\citep{gaertner2002multi}. As we use
the RBF-kernel $k_{rbf}$ as kernel on $\mathcal{X}$ and powers of RBF-kernels
are again RBF-kernels, we will not consider $p$ explicitly in the following.

@@

\subsection{Segment Features}

To describe single segments, we make use of densely computed
SIFT~\citep{lowe2004distinctive} and ColorSIFT~\citep{van2009evaluating}
features, from which we compute bag of visual word histograms. Additionally, we
use histograms of oriented gradients~\citep{dalal2005histograms} on the
segments.  We use RBF-kernels for all of the features, constructing one
MI-kernel per feature. These are then combined using multiple kernel learning
to produce a single kernel matrix. This kernel matrix can then be used for all
classes, making classification particularly efficient.

\subsection{Combining Segments}\seclabel{combiningsegments}

The framework described above yields an image-level and a segment-level
classification.  In our setup, each segment might be given multiple labels.
%TODO explain!
To obtain a pixel-level object-class segmentation, we have to combine these.
When building the segmentation for a given image, we only consider classes
whose presence was predicted on image level.  Since we do not make use of the
ground truth segmentation during training, we cannot learn an optimal
combination as in \citet{li2010object} but perform a simple majority vote
instead.  We merge segments into pixel-level class labels by setting the label
$y_x$ of a pixel $x$ according to
\begin{equation}
    y_x = \argmax_{y \in Y} \#\{S_i | p \in S_i \land y_{S_i}=y \},
\end{equation}
where $Y= \{$car, bike, person$ \}$, $S_i$ enumerates all segments within an
%FIXME we didn't describe the dataset yet!
image and $y_{S_i}$ is the label of segment $S_i$. In words: each pixel is
assigned the class with the highest number of class segments containing it.
This simple heuristic yields good results in practice.

\begin{figure}[tbp]
	\begin{center}
        \includegraphics[width=\linewidth]{images/scheme-crop.pdf}
	\end{center}
        \caption{Overview of our method. See text for details.}
	\figlabel{results}
\end{figure}

\section{Experiments}

\subsection{Instance-Level Predictions using Multi-Instance Kernels}

To assess the validity of instance-level predictions using multi-instance
kernels, we transform $f_{I}$ back to an instance-level classifier, using the
multi-instance learning assumption (\eqref{multi_instance}). We refer to these
instance-based MIK predictions as MIK-instance. In all experiments, the
parameters of the MI-Kernel and SVM are adjusted using MIK and then used with
both MIK and MIK-instance.  This facilitates very fast parameter scans since
MIK is very efficient to compute. Note that we cannot adjust parameters using
instance prediction error, as we assume no instance labels to be known.

@@

\end{table}
We compared the performance of MIK, MIK-instance and state-of-the-art MI
methods on the Musk benchmark datasets~\citep{dietterich1997solving}, see
\Tabref{muskacc}. Results were obtained using 10-fold cross-validation. While
the computational complexity of MIK-instance is very low compared to the other
methods, it achieves competitive results.  Using instance-level labels results
in a slight loss of accuracy of MIK-instances, compared to MIK\@. This small
degradation of performance is quite surprising, since the model was not trained
to provide any instance-level labels.

For multi-class image segmentation, it is beneficial to have a low witness
rate, that is only a few instances are assumed to be positive in a positive
bag. Since an object might not be very prominent in an image, only a fraction
of segments might correspond to the object.
\Tabref{musk-witness} compares the witness rates of MIK-instance,
miSVM~\citep{andrews2003support} and SVR-SVM~\citep{liconvex2010} on the Musk
datasets. MIK-instance is able to achieve similar accuracy with much less
witnesses than the other methods.  Note that Musk1 consists of very small bags
while Musk2 contains significantly larger bags, more similar to the
image/segment setup.

\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccc}
    \toprule
    & \multicolumn2c{Musk1}  & \multicolumn2c{Musk2}  \\
                &accuracy & witness-rate & accuracy & witness-rate  \\
    \cmidrule(rl){2-3}
    \cmidrule(rl){4-5}
    mi-SVM      & 87.4          & 100\%               &  83.6          & 83.9\%\\
    SVM-SVR     & 87.9          & 100\%               &  85.4          & 89.5\%\\
    MIK-instance& 88.0          & 99\%                &  85.2          & 62.3\%\\
    \bottomrule
    \end{tabularx}
    \caption{MIL algorithms and the empirical witness rates of the
        classifiers.  \tablabel{musk-witness}}
\end{table}

@@

\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}p{208pt}ccc}
    \toprule
                & car & bike & person \\
    \cmidrule(l){2-2}
    \cmidrule(l){3-3}
    \cmidrule(l){4-4}
        This work&   0.30&  0.45&  0.43 \\
        Best strongly supervised approaches~\citep{fulkerson2009class,schulz2011}&   0.72&  0.72&  0.66 \\
    \bottomrule
    \end{tabularx}
    \caption{Pixel-level accuracy on the Graz-02 dataset.
        \tablabel{tab:graz}}
\end{table}

Per-class pixel accuracies are reported in \Tabref{tab:graz}; some qualitative
results are shown in \figref{results}. The overall accuracy on images labels,
which is the task that was actually trained, is $90\%$.  While the performance
of our multiple-instance based approach is far from current methods that use
pixel-level annotations, whose pixel-level accuracy is around
$70\%$~\citep{fulkerson2009class,schulz2011} on pixel-level, it can serve as a
baseline for research on weakly supervised methods for object class segmentation.

\section{Summary}

We proposed an algorithm for object-class segmentation using only weak
supervision based on multiple-instance learning. In our approach each image is
represented as a bag of object-like proposal segments.

We described a way to extent bag level predictions made by the multi-instance
kernel method to instance level while remaining competitive with the
state-of-the-art in bag label prediction.

We evaluated the proposed object-class segmentation method on the challenging
Graz02 dataset. While not reaching the performance of methods using strong
supervision, our result can work as a baseline for further research into weakly
supervised object class segmentation.

%In future work, we plan to scale our approach to much larger image datasets. As much
%more images with weak annotations are available than with pixel level segmentation,
%we hope that we can improve upon the state-of-the-art in object-class segmentation
%by making use of larger bodies of training images.
